---
title: "Bis15 Project"
author: "Nick Tobia"
date: "3/10/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library("tidyverse")
```

Read Data
```{r}
data <- readr::read_csv("~/Desktop/BIS15W2020_PBajaj_Project-master/Project/test.csv")
```

Select Out empty rows
```{r}
data2 <- data[-c(1,2,3,4,5,8,20,21,25,28,31,34,35,36,37,38,39,40,41,42,43,50,51,52,53,54,55,56,57,63,64,65,66,67,68,69,70,71,72), ]
```

Select Out Empty Columns
```{r}
df = subset(data2, select = -c(X3,X5,X7,X9,X11,X13,X15,X17,X19) )
```

Renmae Columns
```{r}
data3 <- df %>% 
  rename(Characteristics= "X1", Activity_Reduction="ROLE EMOTIONAL SCALE", Less_Accomplished="X4", Less_Careful_Work= "X6", Physical_Emotional_Health_VS_Social= "SOCIAL FUNCTIONING SCALE", Health_VS_SocialAct.="X10", Nervousness= "MENTAL HEALTH SCALE", Un_Cheerable="X14", Calm="X16",Downhearted="X18", Happy="X20")
```

Separate values into observations and totals.
```{r}
data4 <- data3 %>% 
  separate(Activity_Reduction, into= c("Activity Reduction", "TotalPop"),  sep = "/")  %>% 
  separate(Less_Accomplished, into = c("Less Accomplished", "TotalPop2"), sep = "/")  %>% 
  separate(Less_Careful_Work, into = c("Less Careful Work", "TotalPop3"), sep = "/")  %>%
  separate(Physical_Emotional_Health_VS_Social, into = c("Phys_Emotional Prob Affectting Social Activity", "TotalPop4"), sep = "/")    %>% 
  separate(Health_VS_SocialAct., into = c("Health limits Social Act", "TotalPop5"), sep = "/")   %>% 
  separate(Nervousness, into = c("Nervousness", "TotalPop6"), sep = "/")   %>% 
  separate(Un_Cheerable, into = c("Uncheerable", "TotaPop7"), sep = "/")   %>% 
  separate(Calm, into = c("Calm", "TotalPop8"),sep = "/")      %>% 
  separate(Downhearted, into = c("Downhearted", "TotalPop9"), sep = "/") %>% 
  separate(Happy, into = c("Happy", "TotalPop10"), sep = "/")
data4
```

Remove the percentages from the total columns.
```{r}
data5 <- data4   %>% 
  separate(TotalPop, into = c("TotalPop1", "% Reported1"), sep = " ")  %>% 
  separate(TotalPop2, into = c("TotalPop2", "% Reported2"), sep = " ")  %>% 
  separate(TotalPop3, into = c("TotalPop3", "% Reported3"), sep = " ")  %>% 
  separate(TotalPop4, into = c("TotalPop4", "% Reported4"), sep = " ")  %>% 
  separate(TotalPop5, into = c("TotalPop5", "% Reported5"), sep = " ")  %>% 
  separate(TotalPop6, into = c("TotalPop6", "% Reported6"), sep = " ")  %>% 
  separate(TotaPop7, into = c("TotalPop7", "% Reported7"), sep = " ")  %>% 
  separate(TotalPop8, into = c("TotalPop8", "% Reported8"), sep = " ")  %>% 
  separate(TotalPop9, into = c("TotalPop9", "% Reported9"), sep = " ")  %>% 
  separate(TotalPop10, into = c("TotalPop10", "% Reported10"), sep = " ")
data5
```

```{r}
glimpse(data5)
```


```{r}
 data_subset <- data5 %>% 
  select(Characteristics,`Activity Reduction`, `Less Accomplished`, `Phys_Emotional Prob Affectting Social Activity`, `Nervousness`, `Uncheerable`, `Calm`, `Downhearted`,`Happy` ) %>% 
  filter(Characteristics== "Male" | Characteristics == "Female" |Characteristics == "Leukemia" | Characteristics== "Hodgkin"| Characteristics== "NHL"|Characteristics== "CNS" | Characteristics== "Neuroblastoma" |  Characteristics=="Non-Heritable Retinoblastoma" | Characteristics=="Heritable Retinoblastoma" |Characteristics=="Wilms" | Characteristics== "Bone Sarcoma" | Characteristics == "Soft Tissue Sarcoma" | Characteristics=="Other"| Characteristics=="0-4 years" | Characteristics== "5-9 years" | Characteristics== "10-14 years" | Characteristics=="No" | Characteristics== "Yes" | Characteristics == "Single" | Characteristics=="Married" | Characteristics=="Separated" | Characteristics== "Divorced" | Characteristics== "Widowed" | Characteristics== "Student" | Characteristics== "Never worked/Unemployed" | Characteristics== "Routine/Manual" | Characteristics== "Intermediate" | Characteristics=="Managerial/Professional")
```

Select Male and Female From characteristics
```{r}
plot1 <- data_subset %>% 
  filter(Characteristics=="Male" | Characteristics=="Female") %>% 
  pivot_longer(-Characteristics,
               names_to = "em_state",
               values_to = "count")

plot1$count <- as.numeric(plot1$count)
plot1
```

PLot differences in male and female emotional states as adults 
```{r}
plot1 %>% 
  filter(em_state=="Nervousness" |em_state=="Uncheerable" | em_state=="Calm" | em_state=="Downhearted" | em_state=="Happy"  ) %>%
  ggplot(aes(x=em_state, y=count, group=Characteristics, fill=Characteristics))+
  geom_col(position="dodge")+
  coord_flip()
```

Select Male and Female From characteristics
```{r}
plot2 <- data_subset %>% 
  filter(Characteristics=="Male" | Characteristics=="Female") %>% 
  pivot_longer(-Characteristics,
               names_to = "Soc_State",
               values_to = "Count")

plot2$Count <- as.numeric(plot2$Count)
plot2
```

PLot differences in male and female social and activity states as adults 
```{r}
plot2 %>% 
  filter(Soc_State=="Activity Reduction" |Soc_State=="Less Accomplished" | Soc_State=="Phys_Emotional Prob Affectting Social Activity" ) %>%
  ggplot(aes(x=Soc_State, y=Count, group=Characteristics, fill=Characteristics))+
  geom_col(position="dodge")+
  coord_flip()
```  
Filter cancer as characterisitic
```{r}
plot3 <- data_subset %>% 
  filter(Characteristics=="Leukemia" | Characteristics=="Hodgkin" | Characteristics=="NHL" | Characteristics=="CNS" | Characteristics=="Neuroblastoma" | Characteristics== "Non-Heritable Retinoblastoma" | Characteristics=="Heritable Retinoblastoma" |Characteristics=="Wilms" | Characteristics== "Bone Sarcoma" | Characteristics=="Soft Tissue Sarcoma" | Characteristics== "Other") %>% 
  rename(Diagnosis= "Characteristics")  %>% 
  select(Diagnosis, Nervousness, Uncheerable, Calm, Downhearted, Happy)  %>%
  pivot_longer(-Diagnosis,
               names_to = "Emotional_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))

 
plot3$Observations <- as.numeric(plot3$Observations)
plot3
```

Graph
```{r}
plot3 %>% 
  ggplot(aes(x=Emotional_State, y=Observations, group=Diagnosis, fill=Diagnosis))+
  geom_col(position = "Dodge")+
  ggtitle("Cancer Vs Emotional State")
```

Filter cancer and activity levels
```{r}
plot4 <- data_subset %>% 
  filter(Characteristics=="Leukemia" | Characteristics=="Hodgkin" | Characteristics=="NHL" | Characteristics=="CNS" | Characteristics=="Neuroblastoma" | Characteristics== "Non-Heritable Retinoblastoma" | Characteristics=="Heritable Retinoblastoma" |Characteristics=="Wilms" | Characteristics== "Bone Sarcoma" | Characteristics=="Soft Tissue Sarcoma" | Characteristics== "Other") %>% 
  rename(Diagnosis= "Characteristics")  %>% 
  select(Diagnosis, `Activity Reduction`, `Less Accomplished`, `Phys_Emotional Prob Affectting Social Activity`)  %>%
  pivot_longer(-Diagnosis,
               names_to = "Soc_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))
 
plot4$Observations <- as.numeric(plot4$Observations)
plot4

```

Graph
```{r}
plot4 %>% 
  ggplot(aes(x=Soc_State, y=Observations, group=Diagnosis, fill=Diagnosis))+
  geom_col(position = "Dodge") +
  ggtitle("Cancer Vs Social State") +
  coord_flip() 
```

Separate marriage status and social activities
```{r}
plot5 <- data_subset %>% 
  filter(Characteristics=="Single" | Characteristics=="Married" | Characteristics=="Separated" | Characteristics=="Divorced" | Characteristics=="Widowed") %>% 
  rename(Marriage_Status= "Characteristics")  %>% 
  select(Marriage_Status, `Activity Reduction`, `Less Accomplished`, `Phys_Emotional Prob Affectting Social Activity`)  %>%
  pivot_longer(-Marriage_Status,
               names_to = "Soc_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))
 
plot5$Observations <- as.numeric(plot5$Observations)
plot5

```

Graph
```{r}
plot5 %>% 
  ggplot(aes(x=Soc_State, y=Observations, group=Marriage_Status, fill=Marriage_Status))+
  geom_col(position = "Dodge") +
  ggtitle("Marriage Status Vs Social State") +
  coord_flip() 
```

Select marriage status and emotional state
```{r}
plot6 <- data_subset %>% 
  filter(Characteristics=="Single" | Characteristics=="Married" | Characteristics=="Separated" | Characteristics=="Divorced" | Characteristics=="Widowed") %>% 
  rename(Marriage_Status= "Characteristics")  %>% 
  select(Marriage_Status, Nervousness, Uncheerable, Calm, Downhearted, Happy)  %>%
  pivot_longer(-Marriage_Status,
               names_to = "Emotional_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))

 
plot6$Observations <- as.numeric(plot6$Observations)
plot6
```

Graph
```{r}
plot6 %>% 
  ggplot(aes(x=Emotional_State, y=Observations, group=Marriage_Status, fill=Marriage_Status))+
  geom_col(position = "Dodge")+
  ggtitle("Marriage Status Vs Emotional State")
```

Select ages and emotional state
```{r}
plot7 <- data_subset %>% 
  filter(Characteristics=="0-4 years" | Characteristics=="5-9 years" | Characteristics=="10-14 years") %>% 
  rename(Diagnosis_Age= "Characteristics")  %>% 
  select(Diagnosis_Age, Nervousness, Uncheerable, Calm, Downhearted, Happy)  %>%
  pivot_longer(-Diagnosis_Age,
               names_to = "Emotional_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))

 
plot7$Observations <- as.numeric(plot7$Observations)
plot7
```

Graph
```{r}
plot7 %>% 
  ggplot(aes(x=Emotional_State, y=Observations, group=Diagnosis_Age, fill=Diagnosis_Age))+
  geom_col(position = "Dodge")+
  ggtitle("Diagnosis Age Vs Emotional State")
```

```{r}
plot8 <- data_subset %>% 
  filter(Characteristics=="0-4 years" | Characteristics=="5-9 years" | Characteristics=="10-14 years") %>% 
  rename(Diagnosis_Age= "Characteristics")  %>% 
  select(Diagnosis_Age, `Activity Reduction`, `Less Accomplished`, `Phys_Emotional Prob Affectting Social Activity`)  %>%
  pivot_longer(-Diagnosis_Age,
               names_to = "Soc_State",
               values_to = "Observations") %>%
  arrange(desc(Observations))
 
plot8$Observations <- as.numeric(plot8$Observations)
plot8

```

```{r}
plot8 %>% 
  ggplot(aes(x=Soc_State, y=Observations, group=Diagnosis_Age, fill=Diagnosis_Age))+
  geom_col(position = "Dodge") +
  ggtitle("Diagnosis Age Vs Social State") +
  coord_flip() 
  
```


