---
title: "Lab 4 Homework"
author: "Nick Tobia"
date: "Feb 6 2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove any `#` for included code chunks to run.  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

For this assignment we are going to work with a large dataset from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. First, load the data.  

1. Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.
```{r}
fisheries <- readr::read_csv("data/FAO_1950to2012_111914.csv")
fisheries
```


2. What are the names of the columns? Do you see any potential problems with the column names?

Some of the names are easy to mess up.
Some of the column names are variables, specifically 1950 -> 2012

```{r}
colnames(fisheries)
```


3. What is the structure of the data? Show the classes of each variable.
```{r}
str(fisheries)
```

```{r}
lapply(fisheries,class)
#option glimpse()
```


4. Think about the classes. Will any present problems? Make any changes you think are necessary below.

The Years columns should be numeric, but is currently characters.

```{r}
fisheries <- fisheries %>% 
  mutate_at(vars(starts_with("19")),as.numeric) %>%
  mutate_at(vars(starts_with("2")),as.numeric) %>% 
  mutate_if(is.character,as.factor)

lapply(fisheries, class)
```


5. How many countries are represented in the data? Provide a count.
```{r}
nlevels(fisheries$Country)
```


6. What are the names of the countries?
```{r}
levels(fisheries$Country)
```



7. Use `rename()` to rename the columns and make them easier to use. The new column names should be:  
+ country
+ commname
+ ASFIS_sciname
+ ASFIS_spcode
+ ISSCAAP_spgroup
+ ISSCAAP_spgroupname
+ FAO_area
+ unit
##Reorganized the code to be more readable. Also makes it easier to see inconsistancies in punctuation. Note that the specific errors in this code were identical to errors in anotehr student's code
```{r}
New_fisheries <- (fisheries) %>% 
  rename(country="Country", 
         commname="Common name", 
         ISSCAAP_spgroup="ISSCAAP group#",  
         ISSCAAP_spgroupname="ISSCAAP taxonomic group" , 
         ASFIS_spcode="ASFIS species#", 
         ASFIS_sciname="ASFIS species name", 
         FAO_area="FAO major fishing area", 
         unit="Measure")
names(New_fisheries)
```


8. Are these data tidy? Why or why not, and, how do you know? Use the appropriate pivot function to tidy the data. Remove the NA's. Put the tidy data frame into a new object `fisheries_tidy`.  

##Remember to answer the questions. "Are these data tidy? Why or why not, and, how do you know? "
```{r}
fisheries_tidy <- New_fisheries %>% 
  pivot_longer(-c("country", "commname", "ISSCAAP_spgroup", "ISSCAAP_spgroupname", "ASFIS_spcode", "ASFIS_sciname", "FAO_area","unit"),
               names_to = "Year",
               values_to = "catch",
               values_drop_na = TRUE)
fisheries_tidy
```



9. Refocus the data only to include country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, and catch.
```{r}
fisheries_tidy2 <- fisheries_tidy %>% 
  select(country,ISSCAAP_spgroupname,ASFIS_spcode,ASFIS_sciname, Year, catch)
fisheries_tidy2
```


10. Re-check the classes of `fisheries_tidy2`. Notice that "catch" is shown as a character! This is a problem because we will want to treat it as a numeric. How will you deal with this?
```{r}
lapply(fisheries_tidy2,class)
```

```{r}
fisheries_tidy2$catch <- as.numeric(fisheries_tidy2$catch,Na.RM = T)
class(fisheries_tidy2$catch)
```

11. Based on the ASFIS_spcode, how many distinct taxa were caught as part of these data?
1553 distinct taxa were caught as part of these data.

```{r}
nlevels(fisheries_tidy2$ASFIS_spcode)

#note that n_distinct() give the answer 1551
```


12. Which country had the largest overall catch in the year 2000?
Peru had the largest overall catch in the year 2000 [106250101 fish]

```{r}
fisheries_tidy2 %>% 
  filter(Year=="2000") %>% 
  select(country,Year,catch) %>% 
  group_by(country) %>% 
  summarise(catch = sum(catch, na.rm = TRUE)) %>% 
  arrange(desc(catch,by.group = True))

```



13. Which country caught the most sardines (_Sardina_) between the years 1990-2000?
Morocco caught the most sardines between the years 1990 -> 2000 [5785190 sardines]

```{r}
fisheries_tidy2 %>% 
  select(country,catch,Year,ASFIS_sciname) %>% 
  filter(str_detect(ASFIS_sciname,"Sardina")) %>% 
  filter(Year>= 1990 & Year <= 2000) %>% 
  group_by(country) %>% 
  summarise(catch = sum(catch,na.rm = TRUE)) %>% 
  arrange(desc(catch, by.group = TRUE))
```



14. Which five countries caught the most cephalopods between 2008-2012?
China, Peru, Korea, Japan, and chile caught the most cephalopds between 2008 -> 2012

```{r}
fisheries_tidy2 %>% 
  filter(str_detect(ISSCAAP_spgroupname,"Squids"))%>%
  filter(Year >=2008 & Year <= 2012) %>% 
  group_by(country) %>% 
  summarise(catch=sum(catch,na.rm = TRUE)) %>% 
  arrange(desc(catch, by.group = TRUE))
```


15. Given the top five countries from question 12 above, which species was the highest catch total? The lowest?
Dosidicus gigas	was the highest catch total
Todarodes filippovae was the lowest catch total

```{r}
fisheries_tidy2 %>% 
  filter(str_detect(ISSCAAP_spgroupname,"Squids"))%>%
  filter(Year >=2008 & Year <= 2012) %>% 
  group_by(ASFIS_sciname) %>% 
  summarise(catch_total = sum(catch, na.rm = TRUE)) %>% 
  arrange(desc(catch_total))
```

```{r}
fisheries_tidy2 %>% 
  filter(str_detect(ISSCAAP_spgroupname,"Squids"))%>%
  filter(Year >=2008 & Year <= 2012) %>% 
  group_by(ASFIS_sciname) %>% 
  summarise(catch_total = sum(catch, na.rm = TRUE)) %>% 
  arrange((catch_total))
```


16. In some cases, the taxonomy of the fish being caught is unclear. Make a new data frame called `coastal_fish` that shows the taxonomic composition of "Miscellaneous coastal fishes" within the ISSCAAP_spgroupname column.
##Note if you group by sciname, that output is a little more user friendly
```{r}
coastal_fish <- fisheries_tidy2 %>% 
  filter(str_detect(ISSCAAP_spgroupname,"Miscellaneous coastal fishes"))
coastal_fish
```


17. Use the data to do at least one exploratory analysis of your choice. What can you learn?
I wanted to  see which country caught the most costal fishes
I learned the country was India, which caught 28060848

```{r}
coastal_fish %>% 
  group_by(country) %>% 
  select(country,catch) %>% 
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  arrange(desc(catch,by.group = True))
  
  
```


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   